CHEC Study Sweat Chloride

### 2025-04-16 Angela Metcalf/quantifying_cl_cf_carriers.qmd
NACFC updates

```{r data cleaning}
#| message: false
#| warning: false
# Import from REDCap
unlockREDCap(c(rcon = "Sweat Electrolytes in CF Carriers 24-0691"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
df <- exportRecordsTyped(rcon)
# Format
df <- df %>%
  mutate(
    `Cl Right` = as.numeric(cl_right), `Cl Left` = as.numeric(cl_left),
    `Na Right` = as.numeric(na_right), `Na Left` = as.numeric(na_left)
  )
df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")])
df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")])
# Per Angela, exclude number 15 due to additional mutation
df <- df[-which(df$pid == 15), ]
```

>>>>

```{r data cleaning}
#| message: false
#| warning: false
# Import from REDCap
unlockREDCap(c(rcon = "Sweat Electrolytes in CF Carriers 24-0691"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
df <- exportRecordsTyped(rcon)
# Format
df <- df %>%
  mutate(
    `Cl Right` = as.numeric(cl_right), `Cl Left` = as.numeric(cl_left),
    `Na Right` = as.numeric(na_right), `Na Left` = as.numeric(na_left)
  )
df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")], na.rm = T)
df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")], na.rm = T)
df$Ratio <- df$`Mean Na` / df$`Mean Cl`
# Per Angela, exclude some participants
df <- df[!df$pid %in% c(15, 36), ]
```

# NACFC 2025

```{r results='asis'}
interim <- read.csv("./Data_Clean/Interim reduced data set.csv")
interim$`Chloride >= 30`=interim$Chloride >= 30
summary(tableby(
  Carrier ~ kwt(Average.sodium, "medianq1q3", "iqr") +
    kwt(Chloride, "medianq1q3", "iqr") + kwt(Ratio, "medianq1q3", "iqr") +
    `Chloride >= 30`,
  data = interim
), pfootnote = T)
```

### 2025-04-16 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
NACFC updates

```{r}
ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw()
ggplot(df, aes(x = Diagnosis, y = Na)) +
  geom_boxplot() +
  theme_bw()
ggplot(df, aes(x = Diagnosis, y = Na / Cl)) +
  geom_boxplot() +
  theme_bw()
ggplot(df, aes(x = Diagnosis, y = K)) +
  geom_boxplot() +
  theme_bw()
```

>>>>

```{r}
clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
ratiop <- ggplot(df, aes(x = Diagnosis, y = Na / Cl)) +
  geom_boxplot() +
  theme_bw() +
  ylab("Sweat sodium to chloride ratio")+
  xlab("Modulator")
ratiop
kp <- ggplot(df, aes(x = Diagnosis, y = K)) +
  geom_boxplot() +
  theme_bw()
kp
```

# NACFC 2025

```{r}
```

### 2025-04-16 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
NACFC updates

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(readxl)
library(arsenal)
library(ggpubr)
library(lmerTest)
library(performance)
library(gt)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CFEdith Zemanick/Sweat Variability"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r results='asis'}
# Check if ratio >= 1.2
chec$post_ratio_1.2 <- chec$post_ratio >= 1.2
label(chec$post_ratio_1.2) <- "Post-Modulator Sweat Na:Cl >= 1.2"
summary(tableby(modqual ~ post_ratio + post_ratio_1.2,
  data = chec,
  control = list(
    numeric.stats = c("N", "Nmiss2", "meanCI", "range"),
    cat.stats = c("N", "Nmiss2", "countpct")
  )
), pfootnote = T)
```

@@

```{r}
chec_long %>%
  filter(pre_post == "pre") %>%
  ggplot(., aes(x = ratio, y = cl)) +
  geom_point() +
  ylab("Cl") +
  xlab("Na:Cl") +
  theme_bw() +
  facet_wrap(~modqual)
```

@@

```{r}
chec_long %>%
  filter(pre_post == "post") %>%
  ggplot(., aes(x = ratio, y = cl)) +
  geom_point() +
  ylab("Cl") +
  xlab("Na:Cl") +
  theme_bw() +
  facet_wrap(~modqual)
```

@@

```{r}
plot_df <- chec_long
plot_df$modqual[plot_df$pre_post == "pre"] <- "Pre-Modulator"
plot_df$modqual <- factor(plot_df$modqual,
  levels = c(
    "Pre-Modulator", "Lumacaftor / Ivacaftor", "Tezacaftor / Ivacaftor",
    "Ivacaftor", "Elexacaftor / Tezacaftor / Ivacaftor"
  )
)
ggplot(plot_df, aes(x = modqual, y = ratio)) +
  geom_boxplot() +
  ylab("Na:Cl") +
  theme_bw() +
  theme(axis.title.x = element_blank())
```

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(readxl)
library(arsenal)
library(ggpubr)
library(lmerTest)
library(performance)
library(gtsummary)
library(gt)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r results='asis'}
# Check if ratio >= 1.2
chec$post_ratio_1.2 <- chec$post_ratio >= 1.2
label(chec$post_ratio_1.2) <- "Post-Modulator Sweat Na:Cl >= 1.2"
chec$post_cl <- cut(chec$swcavg, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", ">= 30 and < 60", ">= 60")
)
label(chec$post_cl) <- "Post-Modulator Sweat Cl Group"
summary(tableby(modqual ~ post_cl + post_ratio + post_ratio_1.2,
  data = chec,
  control = list(
    numeric.stats = c("N", "Nmiss2", "meanCI", "range"),
    cat.stats = c("N", "Nmiss2", "countpct")
  )
), pfootnote = T)
```

## Normalized ratio by modulator status and age

We fit a logistic regression model with post-modulator ratio >= 1.2 as the binary outcome. The model with an interaction term was significantly better than the mode with only main effects (p=0.003).

```{r results='asis'}
m1 <- glm(post_ratio_1.2 ~ modqual + age, data = chec, family = "binomial")
m1 %>% tbl_regression(., intercept = T, exponentiate = T)
```

The main effects suggest that people on LUM/IVA or TEZ/IVA were less likely to have a post-modulator ratio >= 1.2 compared to those on ETI, but there was no difference for those on Ivacaftor, after adjusting for age. Also, there was no effect of age on the odds of having a post-modulator ratio >= 1.2.

## Normalized ratio by post-ETI sweat chloride

```{r results='asis'}
summary(tableby(post_cl ~ post_ratio + post_ratio_1.2,
  data = chec[chec$modqual == "Elexacaftor / Tezacaftor / Ivacaftor", ],
  control = list(
    numeric.stats = c("N", "Nmiss2", "meanCI", "range"),
    cat.stats = c("N", "Nmiss2", "countpct")
  )
), pfootnote = T)
```

@@

```{r}
sp_pre <- chec_long %>%
  filter(pre_post == "pre") %>%
  ggplot(., aes(x = ratio, y = cl)) +
  geom_point() +
  geom_hline(yintercept = 60, color = "red") +
  geom_vline(xintercept = 1.2, color = "red") +
  ylab("Cl") +
  xlab("Na:Cl") +
  theme_bw() +
  facet_wrap(~modqual) +
  ggtitle("Pre-Modulator")
sp_pre
```

@@

```{r}
sp_post <- chec_long %>%
  filter(pre_post == "post") %>%
  ggplot(., aes(x = ratio, y = cl)) +
  geom_point() +
  geom_hline(yintercept = 60, color = "red") +
  geom_vline(xintercept = 1.2, color = "red") +
  ylab("Cl") +
  xlab("Na:Cl") +
  theme_bw() +
  facet_wrap(~modqual) +
  ggtitle("Post-Modulator")
sp_post
```

@@

```{r}
plot_df <- chec_long
plot_df$modqual[plot_df$pre_post == "pre"] <- "Pre-Modulator"
plot_df$modqual <- factor(plot_df$modqual,
  levels = c(
    "Pre-Modulator", "Lumacaftor / Ivacaftor", "Tezacaftor / Ivacaftor",
    "Ivacaftor", "Elexacaftor / Tezacaftor / Ivacaftor"
  )
)
boxp <- ggplot(plot_df, aes(x = modqual, y = ratio)) +
  geom_boxplot() +
  ylab("Na:Cl") +
  theme_bw() +
  theme(axis.title.x = element_blank())
boxp
```

## Combined figure

```{r}
boxp / (sp_pre + sp_post)
```
